---
title: "Profile_feedback"
author: "Julian Flowers"
date: "19 June 2017"
output:
  word_document: 
    reference_docx: data/template.docx
---

# Rapid analysis of the profilefeedback mailbox

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE)
```

## Introduction

The `profilefedback` mailbox is accessed via the Fingertips website and is a common means of user communication with the Fingertips team, profile creators, and wider PHE. The method outlined cojld equally be applied to other mailboxes collecting feedback.

```{r, message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(tm)
library(tidytext)



```

In order to download emails is a suitable form for further analysis, we have written a programme for extracting the emails into a text file. For the purposes of analysus we are restricting the fields to:

* Date
* Email sender
* Email recipient
* Subject
* Body

We need to read the text file in to -R- using the "not" separator.

```{r read in emails}

emails<- read_delim("HoFJF2.txt", delim = ";")

head(emails)

emails <- readLines("HoFJF2.txt")
emails[1:20]
emails_test <- data.frame(str_split(emails[3], ";", simplify = FALSE)) %>% mutate_if(is.factor, as.character)
colnames(emails_test) <- "Text"

emails_test <- emails_test %>%
  filter(!str_detect(Text,"[Ee]mail| [Tt]witter| HYPERLINK | cid| www"))

emails_test %>%
  filter(!str_detect(Text, ":")) 

```




c("gold2", "navy", "orchid3")


```{r}
emails <- emails %>% mutate_if(is.factor, as.character) %>% select(-V6)
colnames(emails) <- c("date", "to", "subject", "from", "body")

emails1 <- emails %>% mutate( date = substr(date, 1, 10), date = (lubridate::dmy(date)))

start <- range(emails1$date)[1]
end <- range(emails1$date)[2]


```

This short report covers emails sent between `r start` and `r end`.


## Email frequency

```{r}
options(digits = 2)


tot_mails <- nrow(emails1)

daily_mean <- emails1 %>% group_by(date) %>% tally() %>% summarise(daily_mean = mean(n))
```


Since `r start` the `profilefeedback` mailbox has received `r tot_mails` emails. On average there have been `r round(daily_mean,2)` emails a day since the beginning of 2015. Figure 1 shows daily emails - traffic inceased at the beginning of 2016. 

```{r, fig.cap= "Figure 1: Daily emails"}

emails1 %>%
    group_by(date) %>%
    count() %>%
    ggplot(aes(date, n)) +
    geom_point() +
    geom_line() +
    geom_smooth()



```


Tuesday and Thursday tend to be the busiest days for email traffic coinciding with release days.

```{r, fig.cap= "Figure 2: Emails by month and day of the week"}
library(viridis)

emails1 <- emails1 %>%
    mutate(wkday = lubridate::wday(date, label = TRUE),
           month = lubridate::month(date, label = TRUE), 
           year = lubridate::year(date))

emails1 %>%
    group_by(month, wkday, year) %>%
    count() %>%
    ggplot(aes(month, wkday,  fill= n)) +
    geom_raster() +
    facet_wrap(~year) +
    scale_fill_viridis(name = "No of emails") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5))


```

## Topics



```{r}

emails %>%
    group_by(subject) %>%
    tally(sort = TRUE)

```






```{r}
# df <- emails %>%
#         mutate(parser = gsub("([A-Za-z]+).*", "\\1", `From: (Name)`), 
#                contained = mapply(grepl, pattern=parser, x=Body)) %>%
#         mutate(Body = gsub("\\r","",gsub("\\n","",Body))) %>%
#         mutate(Body = mapply(gsub,pattern=paste0(parser,".*"),replacement="",x=Body))
# 
# sum(df$contained)/nrow(df)



```


```{r la emails, fig.height=10}

emails1 %>%
    group_by(from) %>%
    tally(sort = TRUE) %>%
    filter(stringr::str_detect(from, "gov.uk")) %>%
    separate(from, c("name", "org"), sep = "@") %>%
    separate(org, c("place","gov", "uk"), sep = "\\.", remove = FALSE) %>%
    group_by(place) %>%
    summarise(count = sum(n)) %>%
    ggplot(aes(reorder(place, count), count)) +
    geom_col(fill = "blue") +
    coord_flip() +
    theme(axis.text.y = element_text(size = rel(0.9))) +
    labs(x = "")

```

